### 2021-02-26 output/initial_time_series_plots.Rmd
Add crime and origin categories

---
title: "Untitled"
author: "Matthew P J Ashby"
date: "26/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE)

library(lubridate)
library(readxl)
library(tsibble)
library(tidyverse)
```

```{r load_data, include=FALSE}
calls <- bind_rows(
  read_excel("../../cheshire-calls/N8Data - Submitted Version.xlsx", sheet = 1),
  read_excel("../../cheshire-calls/N8Data - Submitted Version.xlsx", sheet = 2)
) %>% 
  janitor::clean_names()

minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

calls <- calls %>% 
  mutate(
    call_origin = case_when(
      call_origin %in% c(
        "Alarm Company", 
        "Email (to Public Contact Mailbox)",
        "Helpdesk",
        "Public Non Emergency (inc. Door Phones and PCPs)",
        "Single Online Home",
        "Social Media"
      ) ~ "public non-emergency",
      call_origin == "Other Emergency Services (inc. Other Forces)" ~
        "other emergency services",
      call_origin == "Police Generated (inc. Call Sign / Collar Numbers)" ~
        "police generated",
      call_origin == "Unknown Choice List Value" | is.na(call_origin) ~ 
        NA_character_,
      TRUE ~ call_origin
    ),
    incident_type = case_when(
      incident_type %in% c(
        "Assistance to Other Agencies",
        "CRB Use Only",
        "CSI To Be Informed",
        "External System",
        "Found Stolen Vehicle",
        "Message to Pass/OOF Enquiries",
        "PNC Markers",
        "Police Vehicle Recovery",
        "Pre-Planned Events",
        "Stop Search",
        "Training"
      ) ~ "other - admin",
      incident_type %in% c(
        "Alarm - Activation", 
        "Alarm - No Response", 
        "Audible Only Alarm",
        "Insecure Premises",
        "Police Installed Alarm",
        "Unauthorised Encampment"
      ) ~ "alarms",
      incident_type %in% c(
        "Bail Breaches/Wanted Person",
        "Bail/Curfew/Wanted",
        "Prison Licence Recall",
        "Warrant Crown Court"
      ) ~ "warrant/bail issue",
      incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
        "animals",
      incident_type %in% c(
        "Firearms - Non Notifiable Crime",
        "Firearms Involved Crime"
      ) ~ "firearms",
      incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
      incident_type %in% minor_categories ~ "other - minor",
      TRUE ~ incident_type
    )
  )
```

```{r plot function}
covid_plot <- function(data, type, title) {

  stopifnot(is.data.frame(data))
  stopifnot(is.vector(type))
  stopifnot(is.character(title) && length(title))

  data %>% 
    filter(incident_type %in% type) %>% 
    mutate(week = as_date(yearweek(incident_date_time))) %>% 
    count(week) %>% 
    ggplot(aes(x = week, y = n)) +
    geom_point() +
    geom_smooth(formula = y ~ x, method = "loess", span = 0.2) +
    # First UK case
    geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
    # First UK lockdown begins
    geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
    # First UK lockdown ends
    geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
    scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    labs(
      title = title,
      x = NULL,
      y = "weekly incident count"
    ) +
    theme_minimal()

}
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r}
covid_plot(calls, "traffic collision", "Road traffic collisions")
```

```{r}
covid_plot(calls, "ASB", "Anti-social behaviour")
```

```{r}
calls %>% 
  filter(incident_type %in% c("ASB")) %>% 
  filter(!is.na(call_origin)) %>% 
  mutate(week = as_date(yearweek(incident_date_time))) %>% 
  count(call_origin, week) %>% 
  mutate(call_origin = fct_reorder(call_origin, n)) %>% 
  ggplot(aes(x = week, y = n, fill = call_origin)) +
  geom_area(position = "fill") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual") + 
  labs(
    title = "Calls by origin",
    x = NULL,
    y = "proportion of incidents of this type",
    fill = "call origin"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```