### 2021-05-03 output/forecast_origin_plots.Rmd
call_origin_plots

---
title: "forecast_origin_plots"
author: "Nads"
date: "28/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = TRUE)

library(lubridate)
library(tidyverse)
library(tsibble)
```

##By inidivdual Catergory

```{r load data, include=FALSE}
call_forecasts <- read_rds(here::here("output/call_origin_forcasts.Rds"))

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )
```

```{r create chart function, include=FALSE}
forecast_chart <- function (forecasts, types) {

  forecasts %>% 
    filter(call_origin %in% types) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_calls)) +
    geom_point(aes(incident_week, actual_calls, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly count of calls for service",
      fill = "actual calls significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}
```

#Other Emergency Services

```{r OES}
forecast_chart(call_forecasts, "other emergency services")
```

#Police Generated

```{r Police Generated}
forecast_chart(call_forecasts, "police generated")
```

#Public Non-Emergency

```{r Public Non-Emergency}
forecast_chart(call_forecasts, "public non-emergency")
```

#ANPR

```{r ANPR}
forecast_chart(call_forecasts, "ANPR")
```

#999

```{r 999}
forecast_chart(call_forecasts, "999")
```

##By Groups