### 2021-05-18 output/forecast_origin_plots.Rmd
updated forecast

---
title: "Calls for Service During 2021 Compared to Pre-Pandemic Forecast - Call Origin"
author: "Matt Ashby, Nadia Kennar, Reka Solymosi"
date: "28/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = TRUE)

library(lubridate)
library(tidyverse)
library(tsibble)
```

```{r load data, include=FALSE}
call_forecasts <- read_rds(here::here("output/call_origin_forecasts.Rds"))

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )
```

```{r create chart function, include=FALSE}
forecast_chart <- function (forecasts, types) {

  forecasts %>% 
    filter(call_origin %in% types) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_calls)) +
    geom_point(aes(incident_week, actual_calls, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly count of calls for service",
      fill = "actual calls significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}
```

## Other Emergency Services

```{r OES}
forecast_chart(call_forecasts, "other emergency services")
```

No noticeable differnce between actual calls and forecast

## Police Generated

```{r Police Generated}
forecast_chart(call_forecasts, "police generated")
```

No noticeable difference between actual cals and forecast (after, manually setting the lower CI of the forecasts to be zero if it is less than zero in the original forecast))

## Public Non-Emergency

```{r Public Non-Emergency}
forecast_chart(call_forecasts, "public non-emergency")
```

No noticeable difference between actual cals and forecast

## ANPR

```{r ANPR}
forecast_chart(call_forecasts, "ANPR")
```

No noticeable difference between actual calls and forecast - ?

## 999

```{r 999}
forecast_chart(call_forecasts, "999")
```

Calls to 999 received pre-pandemic are higher than those actual actual received during the pandemic. This would be expected with reduced opportunities to commit crimes and therefore less crimes to report. However, there is no clear noticeable trend in the changes as unsure if these changes would be expected?