### 2021-06-09 output/attending_or_not.Rmd

---
title: "Attended"
author: "NK MA RS"
date: "07/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(lubridate)
library(tsibble)
library(tidyverse)

calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()
all_calls_attended_forecast <- read_rds(here::here("output/all_attended_forecasts.Rds"))
by_type_attended_forecast <- read_rds(here::here("output/by_type_attended_forecasts.Rds"))

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )

forecast_chart <- function (forecasts, types) {

  forecasts %>% 
    filter(incident_type_new %in% types) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}

```

Looking at whether or not the proportion of calls attended changes. First visualise:

```{r attendedynvis}

calls$attended <- case_when(calls$attended_flag == 1 ~ "Yes", 
                            calls$attended_flag == 0 ~ "No", 
                            TRUE ~ NA_character_)

calls %>% 
  mutate(week = as_date(yearweek(incident_date_time))) %>% 
  count(attended, week) %>% 
  ggplot(aes(x = week, y = n, fill = attended)) +
  geom_area(position = "fill") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual") + 
  labs(
    title = "Proportion of Concern for Safety Incidents by Attended",
    x = NULL,
    y = "Proportion",
    fill = "Attended (yes or no)"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +    guides(fill=guide_legend(nrow=1,byrow=TRUE))
```

Something happened in 2018... In lockdown we see a spike then a regression back to the mean sort of. Let's try with the forecasting.

Plot just for all calls togethershows that percent of calls which were attended increased over the COVID19 pandemic.

```{r}

all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

```

Now we can break this down by call type:

## Admin

```{r Admin}
forecast_chart(by_type_attended_forecast, "Advice/Directions/Admin")
```

## Alarms

```{r Alarms}
forecast_chart(by_type_attended_forecast, "alarms")
```

## Widlife Concern

```{r Wildlife}
forecast_chart(by_type_attended_forecast, "animals")
```

## ASB

```{r ASB}
forecast_chart(by_type_attended_forecast, "ASB")
```

## Burglary - Commercial

```{r Burglary - Commercial}
forecast_chart(by_type_attended_forecast, "Burglary - Business/Commercial")
```

## Burglary - Residential

```{r Burglary - Residential}
forecast_chart(by_type_attended_forecast, "Burglary - Residential")
```

## Civil Disputes

```{r Civil Disputes}

forecast_chart(by_type_attended_forecast, "Civil Disputes")
```

## Police Safety

```{r Police Safety}

forecast_chart(by_type_attended_forecast, "Code Zero")
```

## Minor Injuries

```{r Minor Injuries}

forecast_chart(by_type_attended_forecast, "Collapse/Illness/Injury/Trapped")
```

## Complaints Against Police

```{r Compaints Against Police}

forecast_chart(by_type_attended_forecast, "Complaints Against Police")
```

## Concern for Safety

```{r Concern For Safety}
forecast_chart(by_type_attended_forecast, "Concern For Safety")
```

## Criminal Damage

```{r Criminal Damage}
forecast_chart(by_type_attended_forecast, "Criminal Damage")
```

## Domestic Incident

```{r Domestic Incident}
forecast_chart(by_type_attended_forecast, "Domestic Incident")
```

## Drug Related Offences

```{r Drug Related Offences}
forecast_chart(by_type_attended_forecast,"Drugs")
```

## Firearms

```{r Firearms}
forecast_chart(by_type_attended_forecast, "firearms")
```

## Fraud

```{r Fraud}
forecast_chart(by_type_attended_forecast,"Fraud")
```

## Hate Incident

```{r Hate Incident}
forecast_chart(by_type_attended_forecast,"Hate Incident")
```

## Highway Disruption

```{r Highway Disruption}
forecast_chart(by_type_attended_forecast,"Highway Disruption")
```

## Missing Person

```{r Missing Person}
forecast_chart(by_type_attended_forecast, "Missing Person")
```

## Other - Admin Calls

```{r Other -Admin Calls}
forecast_chart(by_type_attended_forecast,"other - admin")
```

## Other - minor

```{r Other - minor}
forecast_chart(by_type_attended_forecast, "other - minor")
```

## Police Generated

```{r Police Generated}
forecast_chart(by_type_attended_forecast, "Police Generated")
```

## Potential Breach of the Peace

```{r Breach of Peace}
forecast_chart(by_type_attended_forecast,"Prevent Potential Breach of the Peace")
```

## Property Lost/Found

```{r Property Lost/Found}
forecast_chart(by_type_attended_forecast,"Property Lost/Found")
```

## Road Related Traffic Offence

```{r Road Related Traffic}
forecast_chart(by_type_attended_forecast,"Road Related Traffic Offence")
```

## Robbery

```{r Robbery}
forecast_chart(by_type_attended_forecast, "Robbery")
```

## Sexual Offences

```{r Sexual Offences}
forecast_chart(by_type_attended_forecast, "Sexual Offences")
```

## Shoplifitng

```{r Shoplifting}
forecast_chart(by_type_attended_forecast, "Shoplifting")
```

## Sudden Death

```{r Sudden Death}
forecast_chart(by_type_attended_forecast, "Sudden Death")
```

## Suspicious Activity

```{r Suspicious Activity}
forecast_chart(by_type_attended_forecast, "Suspicious Activity")
```

## Suspicious Package

```{r Suspicious Package}
forecast_chart(by_type_attended_forecast, "Suspicious Package")
```

## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
forecast_chart(by_type_attended_forecast, "Theft From Motor Vehicle")
```

## Theft of Motor Vehicle

```{r Theft of Motor Vehicle}
forecast_chart(by_type_attended_forecast, "Theft Of Motor Vehicle")
```

## Theft Other

```{r Theft Other}
forecast_chart(by_type_attended_forecast, "Theft Other")
```

## Traffic collision

```{r Traffic Collisions}
forecast_chart(by_type_attended_forecast, "traffic collision")
```

## Violence and Harassment Incidents

```{r Violence/Harassment}
forecast_chart(by_type_attended_forecast, "Violence/Harassment")
```

## Warrant and Bail Incidents

```{r Warrant/Bail}
forecast_chart(by_type_attended_forecast, "warrant/bail issue")
```