### 2021-11-01 output/calls_vs_imd.Rmd
updated IMD stuff

---
title: "Calls vs IMD"
author: "NK MA RS"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(fable)
library(lubridate)
library(tsibble)
library(tidyverse)
library(readxl)
library(sf)

# Read and clean data

# excel_sheets("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx")
# imd <- read_xlsx(here::here("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019"))  %>%  janitor::clean_names()

# imd <- read_xlsx("C:\\Users\\maczo\\Desktop\\research\\n8_cov\\File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019") %>%  janitor::clean_names()

imd <- read_xlsx("/Users/reka/Dropbox (The University of Manchester)/N8_covid/analysis/n8_covid/File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019") %>%  janitor::clean_names()

cheshire_lsoas <- st_read("../cheshire_lsoas.geojson")
cheshire_imds <- left_join(cheshire_lsoas, imd, by = c("LSOA11NM" = "lsoa_name_2011"))

# calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()
calls <- read_csv("/Volumes/n8_covid/n8_data_v2.csv.gz") %>%  janitor::clean_names()

minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

# Categorise calls
calls <- mutate(
  calls,
  incident_type_new = case_when(
    incident_type %in% c("Distraction Burglary") ~
      "Burglary - Residential",
    incident_type %in% c(
      "Assistance to Other Agencies",
      "CRB Use Only",
      "CSI To Be Informed",
      "External System",
      "Found Stolen Vehicle",
      "Message to Pass/OOF Enquiries",
      "PNC Markers",
      "Police Vehicle Recovery",
      "Pre-Planned Events",
      "Stop Search",
      "Training"
    ) ~ "other - admin",
    incident_type %in% c(
      "Alarm - Activation", 
      "Alarm - No Response", 
      "Audible Only Alarm",
      "Insecure Premises",
      "Police Installed Alarm",
      "Unauthorised Encampment"
    ) ~ "alarms",
    incident_type %in% c(
      "Bail Breaches/Wanted Person",
      "Bail/Curfew/Wanted",
      "Prison Licence Recall",
      "Warrant Crown Court"
    ) ~ "warrant/bail issue",
    incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
      "animals",
    incident_type %in% c(
      "Firearms - Non Notifiable Crime",
      "Firearms Involved Crime"
    ) ~ "firearms",
    incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
    incident_type %in% minor_categories ~ "other - minor",
    TRUE ~ incident_type
  )
)

cheshire_calls <- calls %>% filter(lsoa %in% unique(cheshire_imds$LSOA11NM))

lsoa_filler <- bind_rows(data.frame(lsoa = unique(cheshire_imds$LSOA11NM),
                          pre_post = "Pre-COVID"), 
                         data.frame(lsoa = unique(cheshire_imds$LSOA11NM),
                          pre_post = "Post-COVID"))

plot_imd <- function(inc_type) {

  type_calls <- cheshire_calls %>% filter(incident_type_new == inc_type) %>% 
    group_by(lsoa) %>% count() %>% 
    left_join(cheshire_imds, ., by = c("LSOA11NM" = "lsoa")) %>% 
    mutate(n = replace_na(n, 0))

  labs_df <- type_calls %>% group_by(index_of_multiple_deprivation_imd_decile) %>% count()

 p <-  ggplot(type_calls %>% filter(!is.na(index_of_multiple_deprivation_imd_decile)), aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
    geom_boxplot() + 
    geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
    scale_y_log10() + 
    labs(title = paste0(inc_type, " calls from LSOAs in each IMD decile 2015-2020")) + 
    xlab("IMD decile (1 = most and 10 = least deprived)") + 
    ylab(paste0("Number of ",inc_type," calls")) + 
    theme_bw()

 return(p)

}

plot_imd_prepost <- function(inc_type) {

  pre_calls <- cheshire_calls %>% 
    filter(incident_type_new == inc_type) %>% 
  mutate(pre_post = ifelse(incident_date_time < dmy("31/01/2020"), "Pre-COVID", "Post-COVID")) %>%   group_by(lsoa, pre_post) %>% count() %>% 
  left_join(lsoa_filler, .) %>%
    mutate(n = replace_na(n, 0)) %>% 
    left_join(., cheshire_imds, by = c("lsoa" = "LSOA11NM")) %>% 
  filter(!is.na(index_of_multiple_deprivation_imd_decile)) %>%  # check why NAs
  mutate(n = replace_na(n, 0), 
         n_per_yr = ifelse(pre_post == "Pre-COVID", n/5,n))

labs_df <- pre_calls %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% summarise(n = n_distinct(lsoa))

 p <-  ggplot() + 
  geom_boxplot(data = pre_calls, aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n_per_yr, color = pre_post)) + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = max(pre_calls$n_per_yr, na.rm = TRUE) + 10, label = paste0("n = ",n))) +
  labs(title = "Total calls from LSOAs in each IMD decile pre and post COVID") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab(paste0("Number of ",inc_type," calls (per year)")) + 
       scale_y_log10() + 
  theme_bw() + 
  theme(legend.position = c(0.87, 0.75))
 return(p)

}

```

**NOTE** there are 17 LSOAs where the IMD ranking isn't there? It's not in the spreadsheet, IDK why so need to check. They are: `r cheshire_imds %>% filter(is.na(index_of_multiple_deprivation_imd_decile)) %>% pull(LSOA11NM)`

## Total calls by IMD decile

```{r}

total_calls <- cheshire_calls %>% group_by(lsoa) %>% count() %>% 
  left_join(., imd, by = c("lsoa" =  "lsoa_name_2011")) %>%
  filter(!is.na(index_of_multiple_deprivation_imd_decile))  # check why NAs

labs_df <- total_calls  %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% count()

ggplot(total_calls , aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
  geom_boxplot() + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
  scale_y_log10() + 
  labs(title = "Total calls from LSOAs in each IMD decile 2015-2020") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Number of calls (all types)") + 
  theme_bw()

```

```{r prepostcovidbox}

pre_calls <- cheshire_calls %>% 
  mutate(pre_post = ifelse(incident_date_time < dmy("31/01/2020"), "Pre-COVID", "Post-COVID")) %>%   group_by(lsoa, pre_post) %>% count() %>% 
  left_join(lsoa_filler, .) %>%
    mutate(n = replace_na(n, 0)) %>% 
  left_join(., cheshire_imds, by = c( "lsoa" = "LSOA11NM")) %>%
  filter(!is.na(index_of_multiple_deprivation_imd_decile)) %>%  # check why NAs
  mutate(n = replace_na(n, 0), 
         n_per_yr = ifelse(pre_post == "Pre-COVID", n/5,n))

labs_df <- pre_calls %>% group_by(pre_post, index_of_multiple_deprivation_imd_decile) %>% count()

ggplot(pre_calls, aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n_per_yr, color = pre_post)) + 
  geom_boxplot() + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = max(pre_calls$n_per_yr + 100), label = paste0("n = ",n))) +
  labs(title = "Total calls from LSOAs in each IMD decile pre and post COVID") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Number of calls (per year)") + 
  theme_bw() + 
  theme(legend.position = c(0.87, 0.75))

```

## Same but by call type

## Admin

```{r Admin}
plot_imd("Advice/Directions/Admin")
plot_imd_prepost("Advice/Directions/Admin")

```

## Alarms

```{r Alarms}
plot_imd("alarms")
plot_imd_prepost("alarms")

```

## Widlife Concern

```{r Wildlife}
plot_imd("animals")
plot_imd_prepost("animals")
```

## ASB

```{r ASB}
plot_imd("ASB")
plot_imd_prepost("ASB")
```

## Burglary - Commercial

```{r Burglary - Commercial}
plot_imd("Burglary - Business/Commercial")
plot_imd_prepost("Burglary - Business/Commercial")
```

## Burglary - Residential

```{r Burglary - Residential}
plot_imd("Burglary - Residential")
plot_imd_prepost("Burglary - Residential")
```

## Civil Disputes

```{r Civil Disputes}

plot_imd("Civil Disputes")
plot_imd_prepost("Civil Disputes")
```

## Police Safety

```{r Police Safety}

plot_imd("Code Zero")
plot_imd_prepost("Code Zero")
```

## Minor Injuries

```{r Minor Injuries}

plot_imd("Collapse/Illness/Injury/Trapped")
plot_imd_prepost("Collapse/Illness/Injury/Trapped")
```

## Complaints Against Police

```{r Compaints Against Police}

plot_imd("Complaints Against Police")
plot_imd_prepost("Complaints Against Police")
```

## Concern for Safety

```{r Concern For Safety}
plot_imd("Concern For Safety")
plot_imd_prepost("Concern For Safety")
```

## Criminal Damage

```{r Criminal Damage}
plot_imd("Criminal Damage")
plot_imd_prepost("Criminal Damage")
```

## Domestic Incident

```{r Domestic Incident}
plot_imd("Domestic Incident")
plot_imd_prepost("Domestic Incident")
```

## Drug Related Offences

```{r Drug Related Offences}
plot_imd("Drugs")
plot_imd_prepost("Drugs")
```

## Firearms

```{r Firearms}
plot_imd("firearms")
plot_imd_prepost("firearms")
```

## Fraud

```{r Fraud}
plot_imd("Fraud")
plot_imd_prepost("Fraud")
```

## Hate Incident

```{r Hate Incident}
plot_imd("Hate Incident")
plot_imd_prepost("Hate Incident")
```

## Highway Disruption

```{r Highway Disruption}
plot_imd("Highway Disruption")
plot_imd_prepost("Highway Disruption")
```

## Missing Person

```{r Missing Person}
plot_imd("Missing Person")
plot_imd_prepost("Missing Person")
```

## Other - Admin Calls

```{r Other -Admin Calls}
plot_imd("other - admin")
plot_imd_prepost("other - admin")
```

## Other - minor

```{r Other - minor}
plot_imd("other - minor")
plot_imd_prepost("other - minor")
```

## Police Generated

```{r Police Generated}
plot_imd("Police Generated")
plot_imd_prepost("Police Generated")
```

## Potential Breach of the Peace

```{r Breach of Peace}
plot_imd("Prevent Potential Breach of the Peace")
plot_imd_prepost("Prevent Potential Breach of the Peace")
```

## Property Lost/Found

```{r Property Lost/Found}
plot_imd("Property Lost/Found")
plot_imd_prepost("Property Lost/Found")
```

## Road Related Traffic Offence

```{r Road Related Traffic}
plot_imd("Road Related Traffic Offence")
plot_imd_prepost("Road Related Traffic Offence")
```

## Robbery

```{r Robbery}
plot_imd("Robbery")
plot_imd_prepost("Robbery")
```

## Sexual Offences

```{r Sexual Offences}
plot_imd("Sexual Offences")
plot_imd_prepost("Sexual Offences")
```

## Shoplifitng

```{r Shoplifting}
plot_imd("Shoplifting")
plot_imd_prepost("Shoplifting")
```

## Sudden Death

```{r Sudden Death}
plot_imd("Sudden Death")
plot_imd_prepost("Sudden Death")
```

## Suspicious Activity

```{r Suspicious Activity}
plot_imd("Suspicious Activity")
plot_imd_prepost("Suspicious Activity")

```

## Suspicious Package

```{r Suspicious Package}
plot_imd("Suspicious Package")
plot_imd_prepost("Suspicious Package")

```

## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
plot_imd("Theft From Motor Vehicle")
plot_imd_prepost("Theft From Motor Vehicle")

```

## Theft of Motor Vehicle

```{r Theft of Motor Vehicle}
plot_imd("Theft Of Motor Vehicle")
plot_imd_prepost("Theft Of Motor Vehicle")

```

## Theft Other

```{r Theft Other}
plot_imd("Theft Other")
  plot_imd_prepost("Theft Other")

```

## Traffic collision

```{r Traffic Collisions}
plot_imd("traffic collision")
plot_imd_prepost("traffic collision")

```

## Violence and Harassment Incidents

```{r Violence/Harassment}
plot_imd("Violence/Harassment")
plot_imd_prepost("Violence/Harassment")

```

## Warrant and Bail Incidents

```{r Warrant/Bail}
plot_imd("warrant/bail issue")
plot_imd_prepost("warrant/bail issue")
```