### 2022-03-14 output/forecast_gini_plots.Rmd
updating n8 presentation

---
title: "Gini Coefficient forecast and observed"
author: "RS MA NK EK"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = TRUE)

library(lubridate)
library(tidyverse)
library(tsibble)
```

```{r load data, include=FALSE}

gini_all_types <- read_rds("/Volumes/n8_covid/gini_forecasts.Rds")
gini_asb <- read_rds("/Volumes/n8_covid/gini_asb_forecast.Rds")
gini_drugs <- read_rds("/Volumes/n8_covid/gini_drugs_forecast.Rds")
gini_dv <- read_rds("/Volumes/n8_covid/gini_dv_forecast.Rds")
gini_violence <- read_rds("/Volumes/n8_covid/gini_violence_forecast.Rds")
gini_misper <- read_rds("/Volumes/n8_covid/gini_misper_forecast.Rds")

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )

make_gini_plot <- function(x){

  x %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_gini)) +
    geom_point(aes(incident_week, actual_gini, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21"))) +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(title = "Gini coefficient* over time",
    subtitle = "*measure of inequality with 0 = perfect equality and 1 = maximal inequality",
      caption = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      x = NULL,
      y = "Gini coefficient",
      fill = "Observed Gini index significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}

```

## Gini for all calls

```{r giniforalltypes}

make_gini_plot(gini_all_types)

```

## Gini forecast for ASB

```{r giniASB}

make_gini_plot(gini_asb )

```

## Gini forecast for drugs

```{r ginidrugs}

make_gini_plot(gini_drugs )

```

## Gini forecast for violence

```{r giniviolence}

make_gini_plot(gini_violence )

```

## Gini forecast for domestic incidents

```{r ginidv}

make_gini_plot(gini_dv)

```

## Gini forecast for missing persons

```{r ginimisper}

make_gini_plot(gini_misper)

```