### 2022-03-21 docs/backup_bc_comments.Rmd
updated slides for N8 webinar

---
title: "Changes to police demand in COVID-19 pandemic"
subtitle: "N8PRP Webinar"
author: "Réka Solymosi, Matt Ashby, Eon Kim, Nadia Kennar, Karen Byrom, Alex McMillan, Phoebe Liebelt"
institute: "University of Manchester, Cheshire Constabulary, UCL"
date: "24/03/2022"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---
class: center, middle

# Aim

Explore changes in calls for service and response over the pandemic and associated lockdowns.

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = TRUE, 
                      message = FALSE, warning = FALSE)

library(lubridate)
library(tidyverse)
library(tsibble)
library(sf)
library(ineq)

all_calls <- read_rds("/Volumes/n8_covid/all_calls_forecast.Rds")
call_forecasts <- read_rds("/Volumes/n8_covid/call_forecasts.Rds")
all_rt_forecast <- read_rds("/Volumes/n8_covid/all_resp_time_forecast.Rds")
by_type_rt_forecast <- read_rds("/Volumes/n8_covid/by_type_rt_forecasts.Rds")
by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/by_grade_rt_forecasts.Rds")
all_calls_attended_forecast <- read_rds("/Volumes/n8_covid/all_attended_forecasts.Rds")
by_type_attended_forecast <- read_rds("/Volumes/n8_covid/by_type_attended_forecasts.Rds")
attended_by_grade_forecast <- read_rds("/Volumes/n8_covid/attended_by_grade_forecasts.Rds")
domestic_by_grade_volume_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_volume_forecasts.Rds")
domestic_by_grade_rt_forecast <- read_rds("/Volumes/n8_covid/domestic_by_grade_rt_forecasts.Rds")
domestic_by_grade_attended_forecasts <- read_rds("/Volumes/n8_covid/domestic_by_grade_attended_forecasts.Rds")
by_grade_tos_forecasts <- read_rds("/Volumes/n8_covid/by_grade_tos_forecasts.Rds")
cheshire_lsoas <- st_read("../cheshire_lsoas.geojson")
lsoa_filler <- bind_rows(data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "Pre-COVID"), 
                         data.frame(lsoa = unique(cheshire_lsoas$LSOA11NM),
                          pre_post = "2020"))
total_calls_by_imd <- read_csv("/Volumes/n8_covid/total_calls_by_imd.csv")
weekly_count_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_count_by_lsoa.csv")
weekly_asb_by_lsoa <- read_csv("/Volumes/n8_covid/weekly_asb_by_lsoa.csv")
# gini_asb <- read_rds("/Volumes/n8_covid/gini_asb_forecast.Rds")

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )

forecast_chart <- function (forecasts, variable, value) {

  # get forecast type to string
  df_name <- deparse(substitute(forecasts))

  # conditionally label y axis
  y_axis_label <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast", 
                                           "domestic_by_grade_rt_forecast") ~ "Median response time (in minutes)", 
                         df_name %in% c("call_forecasts", 
                                        "domestic_by_grade_volume_forecasts") ~ "Weekly count of calls for service", 
                         df_name %in% c("by_type_attended_forecast",
                                        "attended_by_grade_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "Weekly % of calls attended", 
                         df_name %in% c("by_grade_tos_forecasts") ~ "Median time on scene (in minutes)")

  # conditionally find actual value column
  actual_col_name <- case_when(df_name %in% c("by_type_rt_forecast", 
                                           "by_grade_rt_forecast",
                                           "domestic_by_grade_rt_forecast") ~ "actual_rt", 
                         df_name %in% c("call_forecasts",
                                        "domestic_by_grade_volume_forecasts") ~ "actual_calls", 
                         df_name %in% c("by_type_attended_forecast",
                                        "attended_by_grade_forecast", 
                                        "domestic_by_grade_attended_forecasts") ~ "actual_attended", 
                         df_name %in% c("by_grade_tos_forecasts") ~ "actual_tos")

  forecasts %>% 
    filter(eval(as.symbol(variable)) %in% value) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, eval(as.symbol(actual_col_name)))) +
    geom_point(aes(incident_week, eval(as.symbol(actual_col_name)), fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21")))+
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = value,
      subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = y_axis_label,
      fill = "Actual observed value significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

}

```

---
class: center, middle

# The problem

> Binary comparisons of one period to another are common, but can be seriously misleading.

---
.center[![COVID19 and crime article](imgs/cov_news_01.png)]

---
class: center, middle
.center[![COVID19 and crime article](imgs/cov_news_03.png)]

---
.center[![COVID19 and crime article](imgs/cov_news_02.png)]
---

.center[![COVID19 and crime article](imgs/cov_news_04.png)]

---

# Can miss the big picture
.center[![miss picture](imgs/miss_whole_picture.png)]

---
# Important to zoom out
.center[![miss picture 2](imgs/miss_pic_2.png)]

---
# Need to account for trends
.center[![trends](imgs/ignore_trends.png)]

---

# Need to account for noise
.center[![noise](imgs/ignore_noise.png)]

---
class: middle
# Forecasting: a better method

.pull-left[Create a synthetic comparison group by using previous crime data to build a model that forecast how much crime would have occurred in the absence of the pandemic.]
.pull-right[![forecasting](imgs/forecasting_book.png)]

---

# Forecasting Calls for Service

```{r totalcallsfig, fig.width=11,fig.height=7}

# Total calls fig
all_calls %>% 
 mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
  ggplot() +
  # Forecast
  geom_ribbon(
    aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
    na.rm = TRUE,
    alpha = 0.5, 
    fill = "grey80"
  ) +
  geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
  # Dates of interest
  geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
  geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
  # Actual calls
  geom_line(aes(incident_week, actual_calls)) +
  geom_point(aes(incident_week, actual_calls, fill = sig), shape = 21) +
  scale_x_date(date_labels = "%e %b\n%Y", 
               limits = as.Date(c("2020-01-06", "2020-12-21"))) +
  scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
  scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
  labs(
    title = "Change in all calls for service during 2020 \ncompared to pre-pandemic forecast",
    subtitle = str_wrap(
      str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
      80
    ),
    caption = "Forecast calculated using data up to 31 January 2020",
    x = NULL,
    y = "Weekly number of calls",
    fill = "actual calls significantly different from forecast"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom", 
        plot.title = element_text(face = "bold"), 
        text = element_text(size = 20)) +
  xlim(c(min(all_calls$incident_week), ymd("2020-12-14")))

```

---
class: center, middle, inverse

# Results

---
class: center, middle, inverse

# Changes in call types

---

# Quiet roads

```{r quietroadsplots, fig.width=11,fig.height=8}

p1 <- forecast_chart(call_forecasts, "incident_type_new" ,"traffic collision") +
  labs(title = "Traffic Collision",
      subtitle = "",
      caption = "",
      x = NULL,
      y = NULL,
      fill = NULL) +
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new" ,"Highway Disruption")+
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) +
  guides(fill="none")
p3 <- forecast_chart(call_forecasts, "incident_type_new" ,"Road Related Traffic Offence")  +
  labs(subtitle = "",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = NULL,
      fill = "Actual observed value significantly different from forecast")

gridExtra::grid.arrange(p1, p2, p3, ncol = 1)

```

---

# Expressions of concern

```{r asbplot, fig.width=11,fig.height=8}

p1 <- forecast_chart(call_forecasts, "incident_type_new", "ASB") + 
  labs(subtitle = "",
      caption = "",
      x = NULL,
      y = NULL,
      fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new", "Drugs")

gridExtra::grid.arrange(p1, p2, ncol = 1)

```

---
# Expressions of concern

> "ASB calls [are] about COVID breaches being coded as ASB... lots of people were reporting other people for breaching the rules. There were a lot of calls where people were very frustrated and they were complaining about their neighbours and the sort of call that sticks in my mind, which I found quite moving quite upsetting really was when somebody would ring in and say that they weren't able to go and visit their elderly relative who lived maybe a mile away... and their next door neighbour was having a party and all these people were turning up." (interview 1)

--

> "... at the start of it, nobody really knew how it was going to affect them. I think there were a lot of people were quite genuinely scared that they were gonna maybe die from it." (interview 3)

--

> "We do get quite a lot of calls, sometimes about oh, you know the the neighbors are smoking drugs in the back garden, you know, sort of weed, that kind of thing." (interview 8)

---

## Advice/Directions/Admin

```{r Admin, fig.width=11,fig.height=8}

forecast_chart(call_forecasts, "incident_type_new", "Advice/Directions/Admin")
```

---

# Calls for information/ reassurance

> "...in the first part of the first lockdown, people were very confused and very worried about breaking the law and very worried about being arrested for and they would ring up and say, can I do this and can I do that so people were actually asking our advice ..." (interview 1)

--

> "...there are also people calling in because they wanted advice about regulations or were uncertain about what they were allowed to do things..." (interview 3)

--

> "I'd say there's definitely an increase in mental health related incidents that we're seeing and I think people are just really struggling with the isolation of being away from everyone... Sometimes people don't know [who to call] to speak to someone for support so they just ring the police. " (interview 4)

---

# Support for such calls

> "We were directing people to the government website... We were we're not really qualified to advise people on the regulations that actually fit their particular area, because there was an area to area thing at certain times during the lockdown where different local councils and had different idea about the rules, interpreted the rules differently..." (interview 1)

> "Getting quite a few calls from elderly people asking for advice ...  advice on guidelines...  Of course we referred them to the government website because they're better off going something like that's got advice in relation to [COVID regulations] (interview 4)

> "We have an FAQ. That specially for FCC stuff. And most of the answers can be found in there... There was plenty of support here at work. ...we had a very large VDU display on the wall with the latest rules and regulations like a like a PowerPoint running through on a continual on a continual rotation." (interview 1)

---

# Presence of guardians/ absence of targets

```{r guardiansplot1, fig.width=11,fig.height=8}

forecast_chart(call_forecasts, "incident_type_new", "Burglary - Residential")

```

---

# Presence of guardians/ absence of targets

```{r guardiansplot2, fig.width=11,fig.height=7}

p1 <- forecast_chart(call_forecasts, "incident_type_new", "Theft From Motor Vehicle") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new", "Theft Of Motor Vehicle") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p3 <- forecast_chart(call_forecasts, "incident_type_new", "Theft Other")

gridExtra::grid.arrange(p1, p2,p3, ncol = 1)

```

---

# Presence of guardians/ absence of targets

```{r guardiansplot3, fig.width=11,fig.height=7}

p1 <- forecast_chart(call_forecasts, "incident_type_new", "Shoplifting") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p2 <- forecast_chart(call_forecasts, "incident_type_new", "Robbery")

gridExtra::grid.arrange(p1, p2, ncol = 1)

```

---

# Violence and harassment

```{r Violence/Harassment, fig.width=11,fig.height=7}
forecast_chart(call_forecasts, "incident_type_new", "Violence/Harassment")
```

---

# Night time economy

> "Kickoffs, if if I can say that Saturday nights drunken night, yeah, those sort of thing, those sort of calls were a massive decrease." (interview 5)

--

> "A lot of the violence incidents that we deal with, especially the likes of Friday night, Saturday night, weekends people have been out drinking and the clubs and pubs and starting fights wasn't happening because the pubs weren't open." (interview 8)

---

# Online harassment

> "There's definitely more online crime. So obviously people are not able to go out and interact in person, but they can over social media and other avenues, such as phone calls, texts - these definitely seem more... I'd say they changed in the sense that the threats were still being made, but not in person, so they will be made over a text or a phone call, social media..." (interview 4)

---

# Domestic incidents

```{r domesticfig, fig.width=11,fig.height=7}

# domestic fig
forecast_chart(call_forecasts, "incident_type_new", "Domestic Incident")

```

---

# Domestic incidents
This does not change when we break down by grade of call.

```{r domesticbygradefig, fig.width=11,fig.height=7}

# Domestic by grade fig

p1 <- forecast_chart(domestic_by_grade_volume_forecasts,"initial_grade_of_response", "Grade 1")
p2 <- forecast_chart(domestic_by_grade_volume_forecasts,"initial_grade_of_response", "Grade 2") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(domestic_by_grade_volume_forecasts,"initial_grade_of_response", "Grade 3") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p4 <- forecast_chart(domestic_by_grade_volume_forecasts,"initial_grade_of_response", "Grade 4") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")

gridExtra::grid.arrange(p1, p2,p3, p4, ncol = 2)

```

---
class: center, middle
# Domestic incidents: no increase

> "We were expecting an increase in domestic violence and domestic violence is a big issue for any force. And I'm not sure that in the early days we saw an increase. Certainly I didn't notice a dramatic increase or anything like that. (interview 3)

--

> "I don't really remember them being any higher in the lock downs than normal, and there were a lot of people were saying that all domestics have gone up, but I'm just guessing that people were spending so much time with each other forced that they're gonna have arguments. " (interview 6)

--

> "I did expect more domestics and things because people were in their addresses, but it didn't. From my personal experience didn't seem to happen." (interview 7)

--

> "... when we get the calls it's like ... when they've been out on the town up to the pubs having drinks and then they come back and they disagree over something so little. And that's when the domestic start." (interview 5)

---
class: center, middle
# Domestic incidents: couldn't call?

> "I remember a lady who was desperately upset. She was very, very distressed and this must have been as lockdown was easing a bit because her husband had actually gone out and this was the first opportunity she'd had to make the call... that Lady had more than likely gone through all sorts before she could actually found the opportunity to ring it in." (interview 1)

---
class: center, middle
# Domestic incidents: change in type

> "I'd say there's quite a lot of child related incidents to deal with domestic ... You know this parents kept this child too long." (interview 4)

--

> "The other thing as well that we found uhm with the kids as well. People keeping their teenagers inside the house. You know, quite a lot of the kids were getting a bit rebellious and they wanted to go out with their mates and (...) they were having arguments with mum and dad about the fact they couldn't leave the house. There was a a few more of those and you know, sort of." (interview 8)

---
class: center, middle, inverse
# Police response

---
class: center, middle
# Attendance rate

```{r allattendedfig, fig.width=11,fig.height=7}

# % attended fig and response time fig
all_calls_attended_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_attended)) +
    geom_point(aes(incident_week, actual_attended, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
               limits = as.Date(c("2020-01-06", "2020-12-21"))) +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Attendance rate for all calls for service during 2020 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly % of calls attended",
      fill = "actual % significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

```

---
class: center, middle
# Attendance rate

```{r bygradeattendedperc, fig.width=11,fig.height=7}

p1 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 1") + 
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) + 
  guides(fill="none")
p2 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 2") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")
p3 <- forecast_chart(attended_by_grade_forecast,"initial_grade_of_response", "Grade 3") + 
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) + 
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, ncol = 2)

```

---
class: center, middle
# Response time

```{r allresptimefig, fig.width=11,fig.height=8}

all_rt_forecast %>%
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>%
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper),
      na.rm = TRUE,
      alpha = 0.5,
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_rt)) +
    geom_point(aes(incident_week, actual_rt, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Median response time for all calls for service during 2020 compared to pre-pandemic forecast",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "Median response time in minutes",
      fill = "Response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

```

---
class: center, middle
# Response time

```{r bygraderesptime, fig.width=11,fig.height=7}

p1 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 1") +
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) +
  guides(fill="none")
p2 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 2") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p3 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 3") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p4 <- forecast_chart(by_grade_rt_forecast,"initial_grade_of_response", "Grade 4") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

```

---
class: center, middle
# Time on scene

```{r bygradetos, fig.width=11,fig.height=7}

p1 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 1") +
  labs(subtitle = "", caption = "", x = NULL, fill = NULL) +
  guides(fill="none")
p2 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 2") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p3 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 3") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")
p4 <- forecast_chart(by_grade_tos_forecasts,"initial_grade_of_response", "Grade 4") +
  labs(subtitle = "", caption = "", x = NULL, y = NULL, fill = NULL) +
  guides(fill="none")

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

```

---
class: center, middle, inverse
# Change in nature of the demand

---

# Index of Multiple Deprivation

```{r totalimd, fig.width=11,fig.height=7}

labs_df <- total_calls_by_imd  %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% count()

ggplot(total_calls_by_imd , aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
  geom_boxplot(aes(fill = as.factor(index_of_multiple_deprivation_imd_decile)), alpha = 0.8) + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
  scale_fill_brewer(palette = "RdYlBu") + 
  scale_y_log10() + 
  labs(title = "Total calls from LSOAs in each IMD decile 2015-2020") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Number of calls (all types)") + 
  labs(fill = "IMD") + 
  theme_bw()

```

---

# No change for overall calls

```{r imdchange, fig.width=11,fig.height=7}

plot_imds <- function(weekly_count_by_lsoa) {

  firstweek <- min(weekly_count_by_lsoa$inc_wk)
lastweek <- max(weekly_count_by_lsoa$inc_wk)

weekly_count_by_lsoa <- weekly_count_by_lsoa %>% 
  filter(inc_wk > firstweek & inc_wk < lastweek)

num_lsoa_per_decile <- weekly_count_by_lsoa %>% 
  select(lsoa, index_of_multiple_deprivation_imd_decile) %>% 
  distinct() %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% 
  count() %>% 
  rename(num_lsoas = n)

thing <- weekly_count_by_lsoa %>% 
  left_join(., num_lsoa_per_decile, by = c("index_of_multiple_deprivation_imd_decile" = "index_of_multiple_deprivation_imd_decile")) %>% 
  select(inc_wk, n, index_of_multiple_deprivation_imd_decile, num_lsoas)  %>% 
  group_by(inc_wk, index_of_multiple_deprivation_imd_decile, num_lsoas) %>%
  summarise(n = sum(n)) %>%
  mutate(calls_per_lsoa = n/num_lsoas) %>% 
  group_by(inc_wk, index_of_multiple_deprivation_imd_decile) %>%
  summarise(calls_per_lsoa = calls_per_lsoa) %>% 
    mutate(percentage = calls_per_lsoa / sum(calls_per_lsoa))

# Plot
ggplot(thing, aes(x=inc_wk, y=percentage, fill=as.factor(index_of_multiple_deprivation_imd_decile))) + 
    geom_area(alpha= 1 , size=0.001, colour="black") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  geom_label(aes(as.Date("2020-01-23"), 0, label = "1"), fill = "white", colour = "grey20", nudge_y = -0.05) +
  geom_label(aes(as.Date("2020-03-23"), 0, label = "2"), fill = "white", colour = "grey20", nudge_y = -0.05) +
  geom_label(aes(as.Date("2020-06-01"), 0, label = "3"), fill = "white", colour = "grey20", nudge_y = -0.05) +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual", palette = "RdYlBu") + 
  labs(
    title = "Proportion of Calls from different IMD deciles",
    subtitle = "1: First UK Case, 2: Lockdown begins, 3: Lockdown ends",
    x = NULL,
    y = "Proportion",
    fill = "IMD decile"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +    guides(fill=guide_legend(nrow=1,byrow=TRUE))

}

plot_imds(weekly_count_by_lsoa)

```

---

# Increase of calls recorded as ASB (concern over COVID-19 breach) from less deprived areas

```{r asbimdchange, fig.width=11,fig.height=7}

plot_imds(weekly_asb_by_lsoa)

```

---
## Less concentration of ASB calls

```{r giniASB2, fig.width=11,fig.height=7}

make_gini_plot <- function(x){

  weekly_gini <- x %>% 
    group_by(inc_wk) %>% 
    summarise(gini = ineq(n), type = "Gini")

  ggplot(weekly_gini, aes(x = inc_wk, y = gini)) + 
    geom_line() + 
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0.35, label = row), data = dates, colour = "grey20") +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2015-01-01", "2020-12-21")))+
    labs(
      title = "Gini coefficient* over time",
      subtitle = "*measure of inequality with 0 = perfect equality and 1 = maximal inequality",
      x = NULL,
      y = "Gini coefficient") +
    theme_minimal() +
    theme(plot.title = element_text(face = "bold"))
}

make_gini_plot(weekly_asb_by_lsoa)

```

---
class: center, middle, inverse
# Conclusions

---

Highlights:

- Expressions of concern
- Calls for advice
- Violence drop
- No change in domestic violence
- For ASB the caller "profile" changed

For more (quiet roads, absence of potential victims, presence of guardians, changes in police response) please see the full report: []()

---
# Recommendations

- Further spatiotemporal analysis
- Share good practice (e.g. FHQ information to clarify rules and regulations)
- Further and more robust look at changes to domestic incidents

---

# Further training outputs

- [UKDS webinar: an introduction to time series analysis and forecasting](https://ukdataservice.ac.uk/events/an-introduction-to-time-series-analysis-and-forecasting-2/)
- A co-developed series for analysts using R and Python